\LARGE \bf Budgeted Multi-Armed Bandit Models for Sample-Based Grasp Planning in the Presence of Uncertainty

### 2014-12-15 paper.tex
changed to TASE paper v0

\usepackage{tikz,hyperref,graphicx,units,subfig}
\usepackage{benktools}

\usepackage{caption}
\renewcommand{\captionfont}{\footnotesize}
\usepackage{sidecap,wrapfig}
\usepackage[ruled,vlined]{algorithm2e}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\abs}[1]{\lvert#1\rvert} 
\newcommand{\norm}[1]{\lVert#1\rVert}
%\newcommand{\suchthat}{\mid}
\newcommand{\suchthat}{\ \big|\ }
\newcommand{\bd}{\mathbf{d}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\bp}{\mathbf{p}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\by}{\mathbf{y}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\bz}{\mathbf{z}}
\newcommand{\bbf}{\mathbf{f}}
\newcommand{\bzero}{\mathbf{0}}
\newcommand{\bG}{\mathbf{G}}
\newcommand{\bA}{\mathbf{A}}
\newcommand{\bW}{\mathbf{W}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\mX}{\mathcal{X}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\mN}{\mathcal{N}}
\newcommand{\mW}{\mathcal{W}}
\newcommand{\mF}{\mathcal{F}}
\newcommand{\bZ}{\mathbf{Z}}

@@

There are a number of algorithms for developing policies to balance exploration and exploitation.
One algorithm is $\epsilon-$greedy, which is the idea of choosing the arm with the highest empirical expected reward with $1-\epsilon$ probability and choosing a random arm with probability $\epsilon$ \cite{barto1998reinforcement}.
A class of algorithms that have stronger theoretical guarantees are from the Upper Confidence Bound (UCB) family.
UCB algorithms maintain the empirical expected reward based off of pulling each arm multiple times, while also estimating an upper bound on the true expected reward using assumptions on the probability distribution of rewards and number of times each arm has been sampled. 
The algorithm chooses the next option based on Thompson sampling~\cite{agrawal2011analysis} or the Gittins index policy~\cite{weber1992gittins}.
When the rewards come from exponential family distribution, UCB minimizes cumulative regret, which is the sum over each sub-optimal arm of the number of times that arm is pulled times the difference between the true expected reward of an optimal arm and the true expected reward of that sub-optimal arm.
%In practice, when the distribution on the rewards of arms is not known, the empirical methods such as $\epsilon-$greedy have shown to have better performance in some situations \cite{kuleshov}.

In the grasping context, one is only interesting in choosing the best grasp at the end of an exploration cycle.  We can formulate our problem as a   ``budgeted multi-armed bandit model" \cite{madani2004budgeted}.
The budgeted multi-armed bandit model focuses on efficiently exploring the arms until a finite stopping time is reached. 
The objective of the budgeted multi-armed bandit model is to minimize the ``simple regret", which is the difference between the true expected reward of an optimal arm and the true expected reward of the arm pulled at the stopping time.
Thus the budgeted multi-armed bandit model reduces to exploration until the very last time step, at which time one exploitation step is taken.

@@

Since we are in an uncertain environment, we are interested in calculating the expected quality, or $E(Q)$. $E(Q)$ is computed by sampling from our distributions on pose, shape or material properties (friction coefficient or density) and averaging the qualities that are computed. Reducing the time to find the best expected grasp from a set of a large number of grasps is the primary focus of the paper.

\subsection{Line of action}
In an uncertain environment one would not actually know the true $g$, thus we have to work with the trajectory of the gripper. Similar to the work of \cite{christopoulos2007handling}, we assume that each gripper finger approaches along a \textit{line of action}, a 1D curve $\gamma(t)$ with endpoints $a$ and $b$ as seen in Fig. \ref{fig:line_of_action}.
A gripper finger starts at point $a$ and moves towards $b$, we assume $a$ is far enough away to be collision free of the object.
Each gripper contact is defined by a line of action, so we assume the following tuple is provided $\Gamma = ( \gamma_1(\cdot),...,\gamma_m(\cdot) )$, which designates a proposed \textit{grasp plan}.
While we currently assume the gripper moves free of noise, this approach would be applicable to high precision robots such as industrial robots. 
Future work will look at how to efficiently sample when the approach trajectory has noise.

@@

@@

\begin{align}\label{eq:problem_def}
\Gamma^* \in \mbox{argmax}_{\Gamma \in G} E(Q(\Gamma))
\end{align}

@@

In solving Eq. \ref{eq:problem_def}, we want to pick the grasp plan with the highest $E(Q(\Gamma))$ determined via Monte-Carlo Integration.
Due to the non-linear nature of our grasp metric, we currently do not have a way to represent the distribution on grasp quality of $p(Q(\Gamma))$ in closed form, which limits the bandit algorithms available to us.

For our exploration policy, we chose to use a method called \textit{successive elimination}. We maintain a subset of the possible grasp plans that are statistically indistinguishable from the best grasp plan.
During exploration, we sample uniformly from this subset.
This has been shown to have good performance when the number of evaluations is large \cite{bubeck2009pure}.
For each grasp plan, we maintain a $95\%$ confidence interval around the expected quality of the grasp plan.
The width of the confidence interval for a grasp plan $\Gamma_i$ that has been sampled $n_i$ times and the samples have standard deviation $\sigma_i$ is:

\vspace{-2ex}
\begin{align}
C_{ i} := \frac{1.96 \sigma_i}{\sqrt{n_i}}.
\end{align}

Let $\hat{\mu_i}$ be the sample mean of the grasp quality for grasp plan $i$.
Then the confidence interval for grasp plan $i$ is $[\hat{\mu_i} - C_i, \hat{\mu_i} + C_i]$ \cite{caflisch1998monte}.  
After a sample from a grasp plan, we check to see if the confidence interval of the sampled grasp intersects with the confidence interval of the current best grasp plan, and if it does not, we prune that grasp plan from the current active set of  grasp plans. Since we prune grasps that are below a confidence interval we have statistical guarantees unlike earlier approaches to this problem \cite{kehoe2012toward}. In future work, we will look at using other exploration strategies.

\section{Evaluating a Grasp on Gaussian Process Implicit Surface}
 In order to solve our problem definition, we must evaluate $E(Q(\Gamma))$ for a given grasp plan $\Gamma$. We will first discuss how the GPIS is constructed, then which grasp metric $Q$ we chose and lastly proceed into evaluating the expectation $E$ efficiently.

@@

\subsection{Calculating the Expected Grasp Quality}
Given a proposed grasp plan $\Gamma$, the expected grasp quality can be evaluated as follows:

\vspace{-2ex}
\begin{align}\label{eq:shape_sampling}
E(Q(\Gamma)) = \int Q(g|S,\Gamma) p(S) dS
\end{align}

Where $Q(g|S,\Gamma)$ is the grasp quality that is computed on a shape sample drawn from $p(S)$. To compute this we intersect the zero crossing of the level set with the propose grasp plan $\Gamma$ and determine the parameters $g$, this has been the approach taken in previous work  \cite{kehoe2012estimating}, \cite{kehoe2012toward},  \cite{christopoulos2007handling}. See Fig. \ref{fig:shape_samples} for an example of what samples drawn from $p(S)$ induced by GPIS look like. For computational reasons we approximate the integral via Monte-Carlo Integration. We use importance sampling to draw from the distribution induced by GPIS and calculate the following

\begin{align*}
E(Q(\Gamma)) \approx \frac{1}{N} \sum_{i=1}^N Q(g|S_i,\Gamma) , \ \  S_i \sim p(S)
\end{align*}

To compute the above distribution we must draw samples from $p(S)$. In order to draw shape samples from a GPIS, one needs to sample from signed distance function, $sd$, over the joint on all points in the workspace $\mathcal{W}$ or $p(sd(\mathcal{W}))$. Since this is a GPIS, we know the following

@@

We compute the expected grasp quality now as follows:

\vspace{-2ex}
\begin{align}\label{eq:grasp_sampling}
E(Q(\Gamma)) = \frac{1}{N} \sum_{i=1}^N Q(g_i) , \ \ g_i \sim p(g)
\end{align}

@@

\begin{equation}
  \bar{z} 
  =
  \frac
    {\int_{\mathcal{W}}x P(sd(x)<0) dx}
    {\int_{\mathcal{W}}  P(sd(x)<0) dx}
\end{equation}

@@

We consider the problem of selecting the best grasp plan, $\Gamma^*$ out of a set $G$. For our experiments we look at selecting the best grasp out of a size of $|G| = 1000$. In Fig. \ref{fig:simple_regret}, we plotted the simple regret for three of the shapes in our data set averaged over 100 runs and compare it to the Monte-Carlo method that randomly chooses a grasp plan to draw a sample from. We initialize both the Monte-Carlo and bandit technique by sampling each grasp 2 times.  This is to seed the standard deviation. We draw samples from our calculated distributions $p(g)$.  Interestingly, regret is minimized at least an order of magnitude faster than the baseline approach for the shapes we considered, thus motivating the use of sampled observations to choose what to sample next.

\begin{figure*}[ht!]
\centering
\includegraphics[width = 16.5cm, height = 9cm]{figures/Slide09.jpg}
\caption{ \footnotesize Top: Comparison of two sampling methods. Red is the multi-armed bandit method that actively chooses which one to sample and Blue is the Monte-Carlo method. We measure their ability to converge in terms of simple regret averaged over 100 runs. We had them determine the best grasp when $|G|=1000$. In all cases the bandit method converges a least a magnitude faster than the pure Monte-Carlo integration. Bottom: Three objects from our data set that we tested on (Tape, Loofa, Water Bottle), we used a visualization technique described in \cite{mahler2015opt}}
\vspace*{-10pt}
\label{fig:simple_regret}
\end{figure*}

>>>>

\usepackage{tikz,hyperref,graphicx,units,subfig}
\usepackage{benktools}
\usepackage{caption}
\renewcommand{\captionfont}{\footnotesize}
\usepackage{sidecap,wrapfig}
\usepackage[ruled,vlined]{algorithm2e}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\abs}[1]{\lvert#1\rvert} 
\newcommand{\norm}[1]{\lVert#1\rVert}
%\newcommand{\suchthat}{\mid}
\newcommand{\suchthat}{\ \big|\ }
\newcommand{\bd}{\mathbf{d}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\bp}{\mathbf{p}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\by}{\mathbf{y}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\bz}{\mathbf{z}}
\newcommand{\bbf}{\mathbf{f}}
\newcommand{\bzero}{\mathbf{0}}
\newcommand{\bG}{\mathbf{G}}
\newcommand{\bA}{\mathbf{A}}
\newcommand{\bW}{\mathbf{W}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\mX}{\mathcal{X}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\mN}{\mathcal{N}}
\newcommand{\mW}{\mathcal{W}}
\newcommand{\mF}{\mathcal{F}}
\newcommand{\bZ}{\mathbf{Z}}

@@

There are a number of algorithms for developing policies to balance exploration and exploitation.
One algorithm is $\epsilon-$greedy, which is the idea of choosing the arm with the highest empirical expected reward with $1-\epsilon$ probability and choosing a random arm with probability $\epsilon$ \cite{barto1998reinforcement}.
A class of algorithms that have stronger theoretical guarantees are from the Upper Confidence Bound (UCB) family.
UCB algorithms maintain the empirical expected reward based off of pulling each arm multiple times, while also estimating an upper bound on the true expected reward using assumptions on the probability distribution of rewards and number of times each arm has been sampled. 
Traditional bandit algorithms minimize cumulative regret, which is the sum over each sub-optimal arm of the number of times that arm is pulled times the difference between the true expected reward of an optimal arm and the true expected reward of that sub-optimal arm.
%In practice, when the distribution on the rewards of arms is not known, the empirical methods such as $\epsilon-$greedy have shown to have better performance in some situations \cite{kuleshov}.

Different to the normal bandit problem, we do not actually care about cumulative regret because the exploration and exploitation stage is decoupled from each other. For example in medical testing for patients it is important to always administer the best treatment for the given patient. However for the case of product testing in cosmetics it does not matter what products are administer in the testing phase, but only what final product is sold to the consumer.

Best arm identification has a wide variety of literature that largely falls into two camps. One where the algorithm terminates once a fixed confidence interval around the best arm is met and one where a set horizon is given and the algorithm has to choose a arm at that stopping time.

\subsection{Fixed Confidence}
In the fixed confidence setting the forecaster seeks to minimize the simple regret until a fixed confidence threshold is met at which point it terminates. Originally the problem was solved by racing algorithms, were methods used Hoeffding inequalities or the empirical Bernstein inequality to prune arms that were deemed not likely to succeed and used uniform allocation to explore the remaining set \cite{maron1993hoeffding} \cite{mnih2008empirical}. These methods were later extended to include not only the best arm, but return the top m arms \cite{gabillon2012best}.

\subsection{Fixed Budget}
In the fixed budget setting the algorithm is given a stopping time and needs to return the best arm at that stopping time. Audibert el al. demonstrated an algorithm called Successive Rejects that given a budget divides up the time each arm is pull an can return the best arm with near optimal probability depending on the hardness of the problem \cite{audibert2010best}. Other UCB like methods have been proposed as well that measure a confidence gap and then pull the arm with the highest confidence interval \cite{gabillon2012best}. In \cite{bubeck2009pure}, they showed a link between simple regret and cumulative regret that allowed
for the analysis of the existing bandit algorithms like UCB1.

For the case of choosing the best grasp, we are working in the scenario where the robot has a fix amount of time to choose a grasp and thus the fixed budget setting is more appropriate.

@@

Since we are in an uncertain environment, we are interested in calculating the expected quality, or $P(Q>0)$, or the probability of achieving force closure, similar too  \cite{christopoulos2007handling}\cite{kehoe2012toward} . $P(Q>0)$ is computed by sampling from our distributions on pose, shape and material properties (friction coefficient and center of mass) and averaging the qualities that are computed. Reducing the time to find the best expected grasp from a set of a large number of grasps is the primary focus of the paper.

\subsection{Types of Uncertainty}
In this work we consider the following types of uncertainty shape, pose, center of mass and friction coefficient. The probabilistic relationship of all these types of uncertainty can be seen in Fig. \ref{fig:graphical_model}. We model each one of these with a corresponding distribution and demonstrate how MAB algorithms can handle all types of uncertainty.

For shape, we use a GPIS representation as described in Sec. \ref{sec:GP}. For pose we sample the centroid of the GPIS from a Gaussian distribution with user covariance matrix. Center of mass a Gaussian  with mean as the expected center of mass, as defined in Eq. \ref{eq:mass}, assuming uniform mass distribution. Lastly, for friction coefficient we represent it as Gaussian with mean around the expected friction coefficient provided by a user.

\begin{figure}[ht!]
\centering
\includegraphics[width = 6cm, height = 4cm]{figures/Slide8.jpg}
\caption{A graphical model that illustrates the relationship between the different types of uncertainty in an object. Center of Mass uncertainty is dependent on the pose and shape of the object, however friction coefficient is independent of all other types. }
\vspace*{-10pt}
\label{fig:graphical_model}
\end{figure}

\subsection{Line of action}
In an uncertain environment one would not actually know the true $g$, thus we have to work with the trajectory of the gripper. Similar to the work of \cite{christopoulos2007handling}, we assume that each gripper finger approaches along a \textit{line of action}, a 1D curve $\gamma(t)$ with endpoints $a$ and $b$ as seen in Fig. \ref{fig:line_of_action}.
A gripper finger starts at point $a$ and moves towards $b$, we assume $a$ is far enough away to be collision free of the object.
Each gripper contact is defined by a line of action, so we assume the following tuple is provided $\Gamma = ( \gamma_1(\cdot),...,\gamma_m(\cdot) )$, which designates a proposed \textit{grasp plan}.

@@

While we currently assume the gripper moves free of noise, this approach would be applicable to high precision robots such as industrial robots. 
Future work will look at how to efficiently sample when the approach trajectory has noise.

@@

\begin{align}\label{eq:problem_def}
\Gamma^* \in \mbox{argmax}_{\Gamma \in G} P(Q(\Gamma)>0)
\end{align}

@@

In solving Eq. \ref{eq:problem_def}, we want to pick the grasp plan with the highest $P(Q(\Gamma)>0)$ determined via Monte-Carlo Integration.
Since the function to test force closure is boolean, we can model our problem as choosing the best arm for a Beta-Bernoulli Process. Methods with very strong guarantees are now available to us \cite{review}. We now review some of them.
\section{Bandit Algorithms For Best Arm Identification}
\subsection{Thompson Sampling}
Thompson Sampling is a Bayesian method for the multi-armed bandit problem. The main idea behind it is to update the conjugate prior for the distribution of the arm recently pulled. Then sample from the prior distribution on all arms and pull the arm with the highest sample drawn. The full algorithm is shown in Algorithm 1.  Empirically this method has been shown to outperform frequentist methods like UCB \cite{chapelle2011empirical}

\subsection{Upper Confidence Bound (UCB)}
The UCB strategy was the first strategy shown to have asymptotic logarithmic regret \cite{lai1985asymptotically} for distributions with bounded rewards. It basically assumes a frequenist perspective and estimates an empirical confidence bound for each arm based on the observations seen so far. Then it pulls the arm with the highest confidence bound.

\subsection{Gittins Index} 
Gittins index treats the bandit problem as an Markov Decision Process where each lever pull is taking a step along a Markov Chain. While, normally the cost of computing an optimal policy would be exponential in the number of K arms. The independence assumption between each arms allows us to decouple them and solve for a policy that models taking pulls along one arm or selecting a reference arm that represents selecting a different possible value. We computed the Gittins indices using the restart method proposed by Katehakis et al. \cite{katehakis1987multi}.

\begin{algorithm}
 \KwResult{Best Arm, $g^*$ }
 For Beta(1,1) prior: \\
\For{ t=1,2,...}{ 
 Draw $p_{j,t} \sim$ Beta($S_{j,t}+1$,$F_{j,t}+1$) for $j = 1,...,k$\\
 Play $I_t=j$ for $j$ with maximum $p_{j,t}$\\
 Observe reward $X_{I_t,t}$\\
 Update poster:\\
 Set $S_{I_t,t+1} = S_{I_t,t} + X_{I_t,t}$ \\
 Set $F_{I_t,t+1} = F_{I_t,t} + 1 - X_{I_t,t}$\\

 }
 \caption{Thompson Sampling}
\end{algorithm}

\section{Evaluating a Grasp on Gaussian Process Implicit Surface}
 In order to solve our problem definition, we must evaluate $P(Q(\Gamma)>0)$ for a given grasp plan $\Gamma$. We will first discuss how the GPIS is constructed, then which grasp metric $Q$ we chose and lastly proceed into evaluating $P(Q(\Gamma)>0)$ efficiently.

@@

\subsection{Calculating the Probability of Force Closure}
Given a proposed grasp plan $\Gamma$ , we draw samples from the shape distribution $P (S )$, the distribution on center of mass $P (z)$ and the distribution on friction coefficient $P (\mu)$. The distribution on force closure can then be estimated as Beta- Bernoulli Process with shape parameters $\alpha$ and $\beta$. Thus, we can write the probability of force closure as follows

\vspace{-2ex}
\begin{align}\label{eq:shape_sampling}
P(Q(\Gamma|S,\mu,z) > 0) = \frac{\alpha}{\alpha + \beta}
\end{align}

Where $Q(\Gamma|S,\mu,z)$ is the grasp quality that is computed on a shape sample drawn from $p(S)$,$p(z)$,$p(\mu)$. To compute this we intersect the zero crossing of the level set with the propose grasp plan $\Gamma$ and determine the parameters $g$, this has been the approach taken in previous work  \cite{kehoe2012estimating}, \cite{kehoe2012toward},  \cite{christopoulos2007handling}. See Fig. \ref{fig:shape_samples} for an example of what samples drawn from $p(S)$ induced by GPIS look like. For computational reasons we approximate the integral via Monte-Carlo Integration. We use importance sampling to draw from the distribution induced by GPIS and  count the number of sucesses of the grasp $S$ and the number of failures $F$ and calculate the following

\vspace{-2ex}
\begin{align}\label{eq:shape_sampling}
P(Q(\Gamma|S,\mu,z) > 0) \approx \frac{(S + \alpha_0)}{(\alpha_0 + S) +( \beta_0 + F)}
\end{align}

Here $\alpha_0$ and $\beta_0$ correspond to the intial shape parameters. To compute the above distribution we must draw samples from $p(S)$. In order to draw shape samples from a GPIS, one needs to sample from signed distance function, $sd$, over the joint on all points in the workspace $\mathcal{W}$ or $p(sd(\mathcal{W}))$. Since this is a GPIS, we know the following

@@

@@

\label{eq:mass}
\begin{equation}
  \bar{z} 
  =
  \frac
    {\int_{\mathcal{W}}x P(sd(x)<0) dx}
    {\int_{\mathcal{W}}  P(sd(x)<0) dx}
\end{equation}

@@

We consider the problem of selecting the best grasp plan, $\Gamma^*$ out of a set $G$. For our experiments we look at selecting the best grasp out of a size of $|G| = 1000$. In Fig. \ref{fig:simple_regret}, we plotted the simple regret averaged over 100 of the shapes in our data set and compare the different methods (UCB, Thompson, Gittins and the naive random allocation). We initialize both the Monte-Carlo and bandit technique by sampling each grasp 1 time. We draw samples from our calculated distributions $p(g)$.  Interestingly, Gittins and Thompson converge much faster than random and UCB. In Fig. \ref{fig:pulls_per_grasp}, you can see that Gittins and Thompson allocate grasp samples to only the grasps of high quality, thus it is quickly able to ignore the low quality grasps. UCB takes a more conservative approach to sample allocation, which leads to poor performance in the best arm identification problem \cite{bubeck2009pure}.

\begin{figure*}[ht!]
\centering
\includegraphics[width = 16.5cm, height = 9cm]{matlab_figures/simple_regret.eps}
\caption{ \footnotesize Comparison of Simple Regret convergence for the four sequential decision methods (Random, UCB, Thompson, Gittins). Graph is averaged over 100 shapes from the Brown Sillohoute Dataset \cite{brown} with a set $|G|=1000$ for each shape. As you can see the Thompson and Gittins method converge almost a magnitude faster than random allocation. UCB does poorly on simple regret which is expected \cite{best_arm}}
\vspace*{-10pt}
\label{fig:simple_regret}
\end{figure*}

\begin{figure*}[ht!]
\centering
\includegraphics[width = 16.5cm, height = 9cm]{matlab_figures/pulls_per_grasp.eps}
\caption{ \footnotesize Comparison of sample per grasp for the four sequential decision methods (Random, UCB, Thompson, Gittins). Graph is averaged over 100 shapes from the Brown Silhouette Dataset \cite{brown} with a set $|G|=1000$ for each shape. The best grasps are ranked 1 and worst are 1000.  As you can see the Thompson and Gittins method have a tendency to pull grasps with high ranking, while UCB has a more conservative policy \cite{best_arm}}
\vspace*{-10pt}
\label{fig:pulls_per_grasp}
\end{figure*}
